Probabilistic Forecasts in Hierarchical~Time~Series

### 2018-04-02 ProbabilisticCoherence.tex
Changes throughout sections 2,3 and 4 to make things more consistent

We let $\bm{y}_t \in \mathbb{R}^n$ be a vector comprised of observations of all variables in the hierarchy at time $t$,  and $\bm{b}_t \in \mathbb{R}^m$ is a vector comprised of observations of all bottom-level series at time $t$. The hierarchical structure of the data imply the following
\begin{equation}
  \bm{y}_t = \bm{Sb}_t,
\end{equation}
where $\bm{S}$ is an $n \times m$ constant matrix encoding the aggregation constraints.  For example for the hierarchy in Figure~\ref{fig1}, $\bm{y}_t = [y_{Tot,t},y_{A,t}, y_{B,t},y_{C,t},y_{AA,t}, y_{AB,t}, y_{BA,t}, y_{BB,t}]'$, $\bm{b}_t = [y_{AA,t}, y_{AB,t}, y_{BA,t}, y_{BB,t}]'$, $m=4$, $n=7$, and 
$$ 
  \bm{S} = \begin{pmatrix} 
               1& 1 &1 &1  \\ 
               1 &1 & 0 &0 \\   
               0&0  & 1 & 1 \\ 
               & \multicolumn{2}{c}{\bm{I}_4} &   
           \end{pmatrix}, 
$$

where $I_4$ is a $4$-dimension identity matrix.

@@

It is desirable that forecasts, whether point forecasts or probabilistic forecasts, should in some sense respect aggregation constraints.  We follow other authors in describing this property as {\em coherence}.   
%While coherent point forecasts have been discussed many times previously, the definitions of coherence previously given are vague and are not easily extended to the situation of probabilistic forecasting.

We now provide new definitions for coherent forecasts in terms of vector spaces that give a geometric understanding of the problem thus facilitating the development of the reconciliation techniques of section~\ref{sec:reconciliation}.

\begin{definition}[Coherent subspace]
Let an $n$-dimensional time series $\bm{y}_t \in \mathbb{R}^n$ be subject to the linear aggregation constraint $\bm{y}_t = \bm{S}\bm{b}_t$, where $\bm{b}_t \in \mathbb{R}^m$ and $\bm{S}$ is an $n \times m$ constant matrix. The $m$-dimensional subspace $\mathfrak{s}\subset \mathbb{R}^n$ that is spanned by the columns of $\bm{S}$, i.e. $\mathfrak{s}=\mbox{span}(\bm{S})$, is defined as the {\em coherent space}. 
\end{definition}

We also denote the $n\times(n-m)$ orthogonal complement of $\bm{S}$ as $\bm{S}_{\perp}$, where $\mathfrak{s}_\perp=\mbox{span}\left(\bm{S}_{\perp}\right)$ is the nullspace of $\bm{S}$. We also note that when $\bm{S}$ is understood as a linear mapping $\bm{S}:\mathbb{R}^m\rightarrow\mathbb{R}^n$ whose image given by the coherent space $\mathfrak{s}$. This is depicted in Figure~\ref{fig2}.

@@

    \node at (2,3) {$\bm{S}$};

    \node at (0,5) {$\mathbb{R}^m$(domain of $\bm{S}$)};
    \node at (4,5) {$\mathbb{R}^n$(codomain of $\bm{S}$)};
    \node at (4,1.0) {$\mathfrak{s}$};
    %\node[bullet,label=below:$f(x)$] at (4,2.5){};

@@

    \end{tikzpicture}
    \newline
  \end{center}
  \caption{The domain, codomain and image of the mapping $\bm{S}$.}\label{fig2}
\end{figure}

@@

\begin{definition}[Coherent Probabilistic Forecasts]\label{def:cohprob}
  Let $(\mathbb{R}^m, \mathscr{F}^{\mathbb{R}^m}, \nu)$ be a probability triple, where $\mathscr{F}^{\mathbb{R}^m}$ is the usual $\sigma$-algebra on $\mathbb{R}^m$. Let $\breve{\nu}$ be a probability measure on $\mathfrak{s}$ with $\sigma$-algebra $\mathscr{F}^{\mathfrak{s}}$.  Here $\mathscr{F}_{\mathfrak{s}}$ is formed as collection of sets $s(A)$, where $s(A)$ denotes the image of the set $A\in \mathscr{F}$ under the mapping $s(.)$ and the mapping $s:\mathbb{R}^m\rightarrow\mathbb{R}^n$ denotes premultiplication by the matrix $\bm{S}$.  Then $\breve{\nu}$ is said to be a \emph{coherent probability measure} iff
  $$
      \breve{\nu}(s(A)) = \nu(A) \quad \forall  A \in \mathscr{F}^{\mathbb{R}^m},
    $$  
\end{definition}

@@

These definitions of the coherent space $\mathfrak{s}$ and coherent point and probabilistic forecasts are defined in terms of the mapping $\bm{S}$ and may give the impression that the bottom level series play an important role in the definition.  However, alternative definitions could be formed using any set of basis vectors that spans $\mathfrak{s}$. For example, consider the most simple three variable hierarchy where $y_{1,t}=y_{2,t}+y_{3,t}$.  In this case the matrix $\bm{S}$ has columns  $(1,1,0)'$ and $(1,0,1)'$ spanning $\mathfrak{s}$ and the mapping $\bm{S}$ transforms $y_{2,t}$ and $y_{3,t}$ into a coherent vector for the full hierarchy.  However the columns $(1,0,1)'$ and $(0,1,-1)'$ also span $\mathfrak{s}$ and define a mapping that transforms $y_{1,t}$ and $y_{2,t}$ into a coherent vector for the full hierarchy.  In general, since there are an infinite number linear bases for $\mathfrak{s}$ the definitions above could be made in terms of any series and not just the bottom level series.  In general we will call the series (or linear combinations thereof) used in the definitions of coherence as \textit{basis series}.  Unless stated otherwise, we will always consider the bottom level series to be the basis series, since this facilitates comparison with existing approaches in the literature.

@@

%Definitions~\ref{def:cohpoint} and~\ref{def:cohprob} facilitate extension to probabilistic forecast reconciliation which we discuss in the next section. 
To  the best of our knowledge, the only other definition of coherent probabilistic forecasts is given by \citet{BenTaieb2017} who define coherent probabilistic forecasts in terms of convolutions. According to their definition, probabilistic forecasts are coherent when a convolution of forecast distributions of disaggregate series is identical to the forecast distribution of the corresponding aggregate series.  Their definition is consistent with our definition, our reason for providing a different definition is that the geometric understanding of coherence will facilitate our definitions of probabilistic forecast reconciliation to which we now turn our attention.

@@

\begin{definition}\label{def:reconpoint}
  Let $\hat{\bm{y}}_{t+h} \in \mathbb{R}^n$ be any set of incoherent forecasts at time $t+h$, and let 
  \begin{equation}
    \tilde{\bm{y}}_{t+h}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h},
  \end{equation} 
  where $\bm{G}$ is an $m\times n$ matrix. We say $\tilde{\bm{y}}_{t+h}$ is ``reconciled'' with respect to the mapping $g(.)$, where $g:\mathbb{R}^n \rightarrow \mathbb{R}^m$ is defined as premultiplication be $\bm{G}$.
\end{definition}

In definition~\ref{def:reconpoint}  $\bm{S}\bm{G}$ is a projection matrix that maps $\mathbb{R}^n$ onto $\mathfrak{s}$. We note that a similar characterisation of reconciliation has been used in previous studies but where the notation $\bm{P}$ is used in place of $\bm{G}$. In principle $g(.)$ can also be a non-linear mapping, a possibility we consider in Section~\ref{nonlinear} {\color{red}[AP: I WILL ADD THIS LATER]}.

@@

To illustrate further note that the columns of $\bm{S}$ and $\bm{R}$ provide a basis for $\mathbb{R}^n$.  As such any vector of incoherent set of point forecasts $\hat{\bm{y}}_{t+h|t} \in \mathbb{R}^n$, can be expressed in terms of coordinates in the basis defined by $\bm{S}$ and $\bm{R}$. Let $\tilde{\bm{b}}'_{t+h|t}$ and $\tilde{\bm{a}}_{t+h|t}$ be the coordinates corresponding to $\bm{S}$ and $\bm{R}$, respectively. The process of reconciliation involves setting $\tilde{\bm{b}}_{t+h|t}$ to be the values of the reconciled bottom-level series and setting $\tilde{\bm{a}}_{t+h}=\bm{0}$ to ensure coherence. From properties of linear algebra it follows that 
%\[
%(\bm{S} ~ \vdots~ \bm{R})
%(\tilde{\bm{b}}'_{t+h},~ \tilde{\bm{t}}'_{t+h})'
%= \hat{\bm{y}}_{t+h},
%\]
\[
\hat{\bm{y}}_{t+h|t} = (\bm{S} ~ \vdots~ \bm{R})
(\tilde{\bm{b}}'_{t+h|t},~ \tilde{\bm{a}}'_{t+h|t})'= \bm{S}\tilde{\bm{b}}_{t+h|t} +  \bm{R}\tilde{\bm{a}}_{t+h|t},
\]
while setting $\tilde{\bm{a}}_{t+h|t}=\bm{0}$ gives the reconciled point forecast
\[
\tilde{\bm{y}}_{t+h|t} = \bm{S}\tilde{\bm{b}}_{t+h|t}
\]

@@

In order to find $\tilde{\bm{b}}'_{t+h|t}$ we require the inverse $(\bm{S} ~ \vdots~ \bm{R})^{-1}$ which is given by
\begin{equation}
(\bm{S} ~ \vdots~ \bm{R})^{-1} = \begin{pmatrix}
(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot \\ \cdots \\ (\bm{S}'_\bot \bm{R})^{-1}\bm{S}'_\bot
\end{pmatrix}\,,
\end{equation}
where $\bm{S}_{\bot}$ and $\bm{R}_{\bot}$ be the orthogonal complements of $\bm{S}$ and $\bm{R}$
respectively. Thus it follows that $\tilde{\bm{b}}_{t+h}=(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot \hat{\bm{y}}_{t+h}$ and $\tilde{\bm{y}}_{t+h|t}=\bm{S}(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot \hat{\bm{y}}_{t+h|t}$.  Here $(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot$ is equivalent to $\bm{G}$ in Definition~\ref{def:reconpoint}.

@@

\begin{definition} \label{def:reconprob}
  Let $(\mathbb{R}^n, \mathscr{F}^{\mathbb{R}^n}, \hat{\nu})$ be an incoherent probability triple   defined on $\mathbb{R}^n$ associated with $\bm{y}_{t+h}$ conditional on all information up to time $t$.  Let $(\mathbb{R}^m, \mathscr{F}^{\mathbb{R}^m}, \nu)$ be a probability triple defined on $\mathbb{R}^m$. Let $g:\mathbb{R}^n \rightarrow \mathbb{R}^m $ be a linear mapping. Let $\nu$ have the following property:
  \begin{equation}
  \nu(A) = \hat{\nu}(g^{-1}(A)), \quad \forall \quad A \in \mathscr{F}^{\mathbb{R}^m}\,,
  \end{equation}
  where $g^{-1}(A)$ denotes the pre-image of the measurable set $A$, that is the set of all points in $\mathbb{R}^n$ that map to a point in $A$ under the mapping $g(.)$ or $g^{-1}(A):=\left\{\bm{y}\in \mathbb{R}^n:\bm{Gy}\in A \right\}$. We define $\tilde{\nu}$ as the reconciled probability measure of $\hat{\nu}$ with respect to the mapping $g(.)$ if the following holds
  \begin{equation}
  \tilde{\nu}(s(A)) = \nu(A)= \hat{\nu}(\bm{g}^{-1}(A)) \qquad \forall A \in \mathscr{F}^{\mathbb{R}^m},
  \end{equation}
  where $\tilde{\nu}(\cdot)$ is the probability measure on the measure space $(\mathfrak{s}, \mathscr{F}_\mathfrak{s})$.
\end{definition}

>>>>

We let $\bm{y}_t \in \mathbb{R}^n$ be a vector comprised of observations of all variables in the hierarchy at time $t$,  and $\bm{b}_t \in \mathbb{R}^m$ is a vector comprised of observations of all bottom-level series at time $t$. The hierarchical structure of the data imply the following
\begin{equation}
  \bm{y}_t = \bm{Sb}_t,
\end{equation}
where $\bm{S}$ is an $n \times m$ constant matrix that encodes the aggregation constraints.  For example for the hierarchy in Figure~\ref{fig1}, $\bm{y}_t = [y_{Tot,t},y_{A,t}, y_{B,t},y_{C,t},y_{AA,t}, y_{AB,t}, y_{BA,t}, y_{BB,t}]'$, $\bm{b}_t = [y_{AA,t}, y_{AB,t}, y_{BA,t}, y_{BB,t}]'$, $m=4$, $n=7$, and 
$$ 
  \bm{S} = \begin{pmatrix} 
               1& 1 &1 &1  \\ 
               1 &1 & 0 &0 \\   
               0&0  & 1 & 1 \\ 
               & \multicolumn{2}{c}{\bm{I}_4} &   
           \end{pmatrix}, 
$$

where $\bm{I}_4$ is a $4\times 4$ identity matrix.

@@

It is desirable that forecasts, whether point forecasts or probabilistic forecasts, should in some sense respect aggregation constraints.  We follow other authors in using the nomenclature{\em coherence} to describe this property.   
%While coherent point forecasts have been discussed many times previously, the definitions of coherence previously given are vague and are not easily extended to the situation of probabilistic forecasting.

We now provide new definitions for coherent forecasts in terms of vector spaces that give a geometric understanding of the problem thus facilitating the development of the probabilistic forecast reconciliation in section~\ref{sec:reconciliation}.

\begin{definition}[Coherent subspace]\label{def:cohspace}
Let an $n$-dimensional time series $\bm{y}_t \in \mathbb{R}^n$ be subject to the linear aggregation constraint $\bm{y}_t = \bm{S}\bm{b}_t$, where $\bm{b}_t \in \mathbb{R}^m$ and $\bm{S}$ is an $n \times m$ constant matrix. The $m$-dimensional subspace $\mathfrak{s}\subset \mathbb{R}^n$ that is spanned by the columns of $\bm{S}$, i.e. $\mathfrak{s}=\mbox{span}(\bm{S})$, is defined as the {\em coherent space}. 
\end{definition}

We also denote the $n\times(n-m)$ orthogonal complement of $\bm{S}$ as $\bm{S}_{\perp}$, where $\mathfrak{s}_\perp=\mbox{span}\left(\bm{S}_{\perp}\right)$ is the nullspace of $\bm{S}$. Also at times is  will be useful to think of pre-multiplication by $\bm{S}$ as a linear mapping from $\mathbb{R}^m$ to $\mathbb{R}^n$ in which case we use the notation $s(.)$.  Although the codomain of $s(.)$ is $\mathbb{R}^n$ its image is the coherent space $\mathfrak{s}$ as depicted in Figure~\ref{fig2}.

@@

    \node at (2,3) {$s$};

    \node at (0,5) {$\mathbb{R}^m$(domain of $s$)};
    \node at (4,5) {$\mathbb{R}^n$(codomain of $s$)};
    \node at (4.5,1.0) {$\mathfrak{s}$(image of $s$)};
    %\node[bullet,label=below:$f(x)$] at (4,2.5){};

@@

    \end{tikzpicture}
    \newline
  \end{center}
  \caption{The domain, codomain and image of the mapping $s$.}\label{fig2}
\end{figure}

@@

\begin{definition}[Coherent Probabilistic Forecasts]\label{def:cohprob}
  Let $(\mathbb{R}^m, \mathscr{F}^{\mathbb{R}^m}, \nu)$ be a probability triple, where $\mathscr{F}^{\mathbb{R}^m}$ is the usual $\sigma$-algebra on $\mathbb{R}^m$. Let $\breve{\nu}$ be a probability measure on $\mathfrak{s}$ with $\sigma$-algebra $\mathscr{F}^{\mathfrak{s}}$.  Here $\mathscr{F}^{\mathfrak{s}}$ is formed as collection of sets $s(A)$, where $s(A)$ denotes the image of the set $A\in \mathscr{F}^{\mathbb{R}^m}$ under the mapping $s(.)$.  Let the measure $\breve{\nu}$ have the property
  $$
      \breve{\nu}(s(A)) = \nu(A) \quad \forall  A \in \mathscr{F}^{\mathbb{R}^m},
    $$  
  A probabilistic forecast is coherent if uncertainty in $\bm{y_{t+h}}$ conditional on all information up to time $t$ is characterised by the probability triple $(\mathfrak{s},\mathscr{F}^{\mathfrak{s}},\breve{\nu})$.
\end{definition}

@@

These definitions of the coherent space $\mathfrak{s}$ and coherent point and probabilistic forecasts are defined in terms of the mapping $s(.)$ and may give the impression that the bottom level series play an important role in the definition.  However, alternative definitions could be formed using any set of basis vectors that spans $\mathfrak{s}$. For example, consider the most simple three variable hierarchy where $y_{1,t}=y_{2,t}+y_{3,t}$.  In this case the matrix $\bm{S}$ has columns  $(1,1,0)'$ and $(1,0,1)'$ spanning $\mathfrak{s}$ and the mapping $\bm{S}$ transforms $y_{2,t}$ and $y_{3,t}$ into a coherent vector for the full hierarchy.  However the columns $(1,0,1)'$ and $(0,1,-1)'$ also span $\mathfrak{s}$ and define a mapping that transforms $y_{1,t}$ and $y_{2,t}$ into a coherent vector for the full hierarchy.  In general, since there are an infinite number linear bases for $\mathfrak{s}$ the definitions above could be made in terms of any series and not just the bottom level series.  In general we will call the series (or linear combinations thereof) used in the definitions of coherence as \textit{basis series}.  Unless stated otherwise, we will always consider the basis series to be the bottom level series as in Definition~\ref{def:cohpoint} and Defintion~\ref{def:cohprob}, since this facilitates comparison with existing approaches in the literature.

@@

%Definitions~\ref{def:cohpoint} and~\ref{def:cohprob} facilitate extension to probabilistic forecast reconciliation which we discuss in the next section. 
To  the best of our knowledge, the only other definition of coherent probabilistic forecasts is given by \citet{BenTaieb2017} who define coherent probabilistic forecasts in terms of convolutions. According to their definition, probabilistic forecasts are coherent when a convolution of forecast distributions of disaggregate series is identical to the forecast distribution of the corresponding aggregate series.  Their definition is consistent with our definition, our reason for providing a different definition is that the geometric understanding of coherence will facilitate our definitions of point and probabilistic forecast reconciliation to which we now turn our attention.

@@

  Let $\hat{\bm{y}}_{t+h|t} \in \mathbb{R}^n$ be any set of incoherent forecasts at time $t+h$ using information up to and including time $t$.  Let $\bm{G}$ be an $m\times n$ matrix and $g:\mathbb{R}^n \rightarrow \mathbb{R}^m$ be the mapping corresponding to pre-multiplication by $\bm{G}$.
  \begin{definition}\label{def:reconpoint}
  The point forecast $\tilde{\bm{y}}_{t+h|t}$ ``reconciles'' $\hat{\bm{y}}_{t+h|t}$ with respect to the mapping $g(.)$ iff
  \begin{equation}
    \tilde{\bm{y}}_{t+h|t}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}\,.
  \end{equation}

\end{definition}

In definition~\ref{def:reconpoint}  $\bm{S}\bm{G}$ is a projection matrix that maps $\mathbb{R}^n$ onto $\mathfrak{s}$. We note that a similar characterisation of reconciliation has been used in previous studies but where the notation $\bm{P}$ is used in place of $\bm{G}$. In principle the definition can be generalised where $g(.)$ is a non-linear mapping, a possibility we consider in Section~\ref{nonlinear} {\color{red}[AP: I WILL ADD THIS LATER]}.

@@

To illustrate further note that the columns of $\bm{S}$ and $\bm{R}$ provide a basis for $\mathbb{R}^n$.  As such any incoherent set of point forecasts $\hat{\bm{y}}_{t+h|t} \in \mathbb{R}^n$, can be expressed in terms of coordinates in the basis defined by $\bm{S}$ and $\bm{R}$. Let $\tilde{\bm{b}}'_{t+h|t}$ and $\tilde{\bm{a}}_{t+h|t}$ be the coordinates corresponding to $\bm{S}$ and $\bm{R}$, respectively. The process of reconciliation involves setting $\tilde{\bm{b}}_{t+h|t}$ to be the values of the reconciled bottom-level series and setting $\tilde{\bm{a}}_{t+h}=\bm{0}$ to ensure coherence. From properties of linear algebra it follows that 
%\[
%(\bm{S} ~ \vdots~ \bm{R})
%(\tilde{\bm{b}}'_{t+h},~ \tilde{\bm{t}}'_{t+h})'
%= \hat{\bm{y}}_{t+h},
%\]
\[
\hat{\bm{y}}_{t+h|t} = (\bm{S} ~ \bm{R})
\begin{pmatrix}
\tilde{\bm{b}}_{t+h|t}\\ \tilde{\bm{a}}_{t+h|t}
\end{pmatrix}= \bm{S}\tilde{\bm{b}}_{t+h|t} +  \bm{R}\tilde{\bm{a}}_{t+h|t},
\]
while setting $\tilde{\bm{a}}_{t+h|t}=\bm{0}$ gives the reconciled point forecast
\[
\tilde{\bm{y}}_{t+h|t} = \bm{S}\tilde{\bm{b}}_{t+h|t}
\]

@@

In order to find $\tilde{\bm{b}}_{t+h|t}$ we require the inverse $(\bm{S} ~ \bm{R})^{-1}$ which is given by
\begin{equation}
(\bm{S} ~  \bm{R})^{-1} = \begin{pmatrix}
(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot \\ (\bm{S}'_\bot \bm{R})^{-1}\bm{S}'_\bot
\end{pmatrix}\,,
\end{equation}
where $\bm{S}_{\bot}$ and $\bm{R}_{\bot}$ be the orthogonal complements of $\bm{S}$ and $\bm{R}$
respectively. Thus it follows that $\tilde{\bm{b}}_{t+h}=(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot \hat{\bm{y}}_{t+h}$ and $\tilde{\bm{y}}_{t+h|t}=\bm{S}(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot \hat{\bm{y}}_{t+h|t}$.  Here $(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot$ is equivalent to $\bm{G}$ in Definition~\ref{def:reconpoint}.

@@

  Let $(\mathbb{R}^n, \mathscr{F}^{\mathbb{R}^n}, \hat{\nu})$ be an incoherent probability triple that characterises forecast uncertainty in for all variables in the hierarchy at time $t+h$ conditional on all information up to time $t$.  Let $g:\mathbb{R}^n \rightarrow \mathbb{R}^m $ be a linear mapping.  Let $(\mathbb{R}^m, \mathscr{F}^{\mathbb{R}^m}, \nu)$ be a probability triple defined on $\mathbb{R}^m$ where $\nu$ has the following property:
  \begin{equation}
  \nu(A) = \hat{\nu}(g^{-1}(A)), \quad \forall \quad A \in \mathscr{F}^{\mathbb{R}^m}\,.
  \end{equation}
  Here $g^{-1}(A)$ denotes the pre-image of the measurable set $A$, that is $g^{-1}(A):=\left\{\bm{y}\in \mathbb{R}^n:\bm{Gy}\in A \right\}$. In words, the preimage $g^{-1}(A)$ is the set of all points in $\mathbb{R}^n$ that $g(.)$ maps to a point in $A$.
  \begin{definition} \label{def:reconprob}
  We define $\tilde{\nu}$ as the reconciled probability measure of $\hat{\nu}$ with respect to the mapping $g(.)$ if the following holds
  \begin{equation}
  \tilde{\nu}(s(A)) = \nu(A)= \hat{\nu}(g^{-1}(A)) \qquad \forall A \in \mathscr{F}^{\mathbb{R}^m},
  \end{equation}
  where $\tilde{\nu}(\cdot)$ is a probability measure on the $\mathfrak{s}$ with $\sigma$-algebra $\mathscr{F}_\mathfrak{s}$.
\end{definition}
